Radial and rotational velocities of young brown dwarfs and very low-mass stars in the Upper Scorpius OB association and the $\rho$~Ophiuchi cloud core

### 2006-08-19 BD_radvel/mn060888mj_rev/kurosawa.tex
RK's revision finished.
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@786 25464f21-6099-402c-b30a-640f1d40d38b

The data were reduced via the standard ESO pipeline procedures for
UVES echelle spectra. In summary, the data were corrected for bias,
interorder background, sky background, sky emission lines and cosmic
ray hits. They were then flattened, optimally extracted, and finally
the different orders were merged. No binning was performed to achieve
high resolution required for the RV measurements. The wavelength was
calibrated using the Thorium-Argon arc spectra with a typical value
of the standard deviation of the dispersion solution of 5~m\AA\,
which corresponds to 0.2~$\kmps$ at the central wavelength 8600~\AA.
However, the autoguiding of the telescope keeps the star at the centre
of the slit with about a tenth of the FWHM ($1\,\kmps$) which sets
the upper limit for the systematic error in the RV measurements \citep{bailer-jones:2004}.
The accuracy of the wavelength calibration will be demonstrated in
the RV measurements of a RV standard in the following section. A typical
signal-to-noise ratio (S/N) per wavelength bin of the spectra is about
15, and the heliocentric velocity correction was applied to the final
spectra.

@@

We have computed $p$ for all the objects (Table~\ref{tab:chisq_prob}),
and have plotted the results  
as a histogram of $-\log{p}$, shown in Fig.~\ref{fig:chisq-test}
(excluding the non-member USco~121; see explanation later). 
The figure clearly shows two distinctive populations: one on the left
(with small $-\log{p}$ values) occupied by the RV constant objects,
and one on the right (with the large $-\log{p}$ values)
occupied by the RV variable candidates.  The
expected distribution of $p$ computed consistently with our
uncertainty measurements is also shown in the same figure. The RV
constant population on the left side reasonably matches the expected
curve. 
\textcolor{red}{To quantify this point, we have computed the
$\chi^{2}$ probability to test the goodness of the fit of the expected
distribution to the histogram. We restrict the fit to the first 4 bins
from the left $\left(0 < -\log_{10}\left(p\right) < 2 \right)$ the
histogram since the objects in the second bin from the right
$\left(2.5 < -\log_{10}\left(p\right) < 3 \right)$ are `potentially'
RV-variable object, although we have flagged them as the RV-constant
objects. In this analysis, we find the probability of 0.5 which
indicates that the fit is reasonable. Further, if we re-normalized the
expected distribution, accounting for that fact we neglected the
objects in the second bin from the right, we find the $\chi^{2}$
probability of the fit increases to 0.9.  This, in turn, indicates}
that our uncertainties in RV values are reasonable, and an additional
systematic error may not be necessary in this analysis.

@@

\caption{Histogram of the average heliocentric radial velocities of 14 UScoOB BD and VLM
objects listed in Table~\ref{tab:summary_results} (excluding USco~121,
a non-member). The gaussian fit (dashed) of the radial velociity
distribution gives a standard deviation of $1.0\,\kmps$ and the peak
position of $-6.3\,\kmps$. }
\label{fig:vrad_histrogram}

@@

The range of $v\,\sin i$ found among our objects is $3.6$--$55.6\,\kmps$,
and a similar range is also found by \citet{mohanty:2005}. 
Fig.~\ref{fig:vrot_histrogram} shows the histogram
of $v\,\sin i$ distribution for the UScoOB objects (14 objects excluding
USco 121, non member). The log-normal fit of this distribution gives
the peak position at $16.9\,\kmps$ with a standard deviation $\sigma=27.8\,\kmps$.
Using the $v\,\sin i$ data in \citet{joergens:2001}, the same histogram
bin size used for UScoOB obects and a log-normal fit, we find the
$v\,\sin i$ distribution of the BD and VLM stars (8 objects) in Cha~I
peaks at $15.4\,\kmps$, and has the standard deviation of $8.0\,\kmps$.
The peak of the distribution is similar to that of Upper Sco objects,
but the standard deviation of the distribution is much smaller
than that of our Upper Sco objects. The difference maybe due to the
very small sample. A similar fit was applied to the $v\,\sin i$
distribution of 14 higher mass T~Tauri stars in Cha~I using the data
of \citet{joergens:2001}, and we found a peak at $17.0\,\kmps$ with a
standard deviation $25.9\,\kmps$ which are very similar to those of
the Upper Sco brown dwarf candidates and VLM stars.

@@

\caption{Histogram of the rotational velocities ($v\,\sin{i}$) of 14 UScoOB BD and VLM
objects listed in Table~\ref{tab:summary_results} (excluding USco~121,
a non-member).  The log-normal fit (dashed) of the
rotational velocity distribution gives a standard deviation
$27.8\,\kmps$ and the peak poistion $16.9\,\kmps$.}
\label{fig:vrot_histrogram}

@@

In order to compute the detection probability as a function of
semimajor axis $a$, we take the following procedure: (1)~for each
object in our targets, we randomly select $10^5$ binaries using the
assumption stated above for a given value of $a$, and compute the RV
of the primary ($V_{1}$), (2)~evolve the orbit by the time separation
of the RV measuremets used in the observations for this object, and
take another simulated measurement of RV ($V_{2}$), (3)~from $V_{1}$,
$V_{2}$ and the average uncertainty in RV ($\sigma_{\mathrm{RRV}}$)
from the observations, we compute the $\chi^{2}$ probability $p$,  and
flag the trial as a detection if $p<10^{-3}$ as done for the real
data, (4)~find the fraction of detections out of all random trials,
(5) repeat 1--4 for the range of $a$ between $10^{-3}$~and 10~au, and
(6) repeat 5 for the all targets, and find the detection probability
averaged over all targets as a function of $a$.

@@

In Section~\ref{sub:Radial-velocities}
(see Fig.~\ref{fig:chisq-test}), we found the histogram of the
$\chi^{2}$ probability $p$, for the RV data fit assuming constant RV,
reasonably matches the expected distribution.  This leads us to
conclude that the uncertainty estimates in RV values are 
reasonable.  There are two possible
source of errors: (1)~the RV shifts caused by surface/atmospheric
activity, and (2)~the RV shifts due to the limit of the stability of
autoguiding of the telescope.  \citet{joergens:2006} found that (1) is
negligibly small for young VLM objects with $M_{*}<0.12\,M_{\sun}$.
The size of RV errors caused by (2) is expected to be $<1\,\kmps$
\citep{bailer-jones:2004}. Note that \citet{maxted:2005} found an
additional systematic error was not needed to be added to the RV
uncertainties of \citet{joergens:2006} who used the same instruments
as ours, and used the same method of estimating the
uncertainties. Hence, we expect the systematic error to be much
smaller than $1\,\kmps$ if one is required.

An alternative explanation for the RV variations found in the binary
candidates above is a type of stellar pulsations found by
\citet{palla:2005}. They studied the non-adiabatic, linear
instability of very VLM stars and BD during the deturium burning phase
in the core, and found unstable fundamental modes in the time-scale
between $\sim 1$ and $\sim 5$~h for the object mass from $0.02$ and
$0.1\,\Msun$.  By using their pulsation periods and by assuming that the
pulsation amplitudes is 5~per~cent of the radius of the objects, we
find that the expected RV variation can be $>6\,\kmps$.
As mentioned in Section~\ref{sec:observation-reduction}, for each object at a
given night, two separate spectra are obtained consecutively (with
time separations of $\sim 0.5$~h.). We have measured the RV from each
spectrum to check if there is a large jump in the RV measurements on a
short time-scale. The size of the shifts in the RV values measured in
the two consecutive spectra are found in between $0.02\pm0.10$
(USco~128) and  $2.04\pm1.67\,\kmps$ (USco~66), and the average size
of the shifts $0.36\,\kmps$, which is much smaller than the 
expected RV variation based on the models of \citet{palla:2005};
hence, we found little indication of the pulsations in our sample.

@@

Follow-up spectroscopic observations of the binary candidates
presented here are planned in near future. There are only a few RV
variable binary candidates identified in earlier surveys
(\citealt{guenther:2003}; \citealt{kenyon:2005}; \citealt{joergens:2006}).
As \citet{burgasser:2006} points out most of the current RV and
imaging surveys use samples from magnitude-limited survey, but one
should attempt to use the samples from volume-limited survey in order
to find a correct statistics on binary parameters more straightforwardly,
i.e. without correcting for bias.

@@

We thank the staff of VLT of the ESO for carrying out the observations
in service mode. RK is grateful for Rob Jeffries  and Tim Naylor for helpful
suggestions on the data analysis presented in this paper. We also
thank Matthew Bate for providing us valuable comments on
the manuscript. This work was supported by PPARC rolling grant
PP/C501609/1.

@@

\bibitem[{{Basri} \& {Reiners}(2006)}]{basri:2006}
{Basri} G., {Reiners} A., 2006, preprint (astro-ph/0604259)

@@

% LocalWords:  BDs VLM OB Ophiuchi al Ryuichi Harries QL RV UScoOB BD echelle
% LocalWords:  VLT UVES RVs Cha bouy burgasser gizis kraus basri maxted PPl EEV
% LocalWords:  pinfield substellar duquennoy VLMs guenther joergens padoan CCDs
% LocalWords:  reipurth whitworth dezeeuw ardila luhman muzerolle Kueyen Parnal
% LocalWords:  LL ESO interorder autoguiding FWHM jones Oph wilking mohanty Sp
% LocalWords:  GY USco LHS HJD gaussian Cen garcia sanchez HD RRV pc CCF exp
% LocalWords:  obs

>>>>

The data were reduced via the standard ESO pipeline procedures for
UVES echelle spectra. In summary, the data were corrected for bias,
interorder background, sky background, sky emission lines and cosmic
ray hits. They were then flattened, optimally extracted, and finally
the different orders were merged. No binning was performed to achieve
high resolution required for the RV measurements. The wavelength was
calibrated using the Thorium-Argon arc spectra with a typical value
of the standard deviation of the dispersion solution of 5~m\AA\,
which corresponds to 0.2~$\kmps$ at the central wavelength 8600~\AA.
However, the autoguiding of the telescope keeps the star at the centre
of the slit with about a tenth of the FWHM ($1\,\kmps$) which sets
the upper limit for the systematic error in the RV measurements
\citep{bailer-jones:2004}.
\textcolor{red}{
As we will find out in the following section
(Section~\ref{sub:Radial-velocities}), the systematic error may be
ignored based on result of the $\chi^{2}$ probability analysis we
performed without the introduction of assumed systematic error.
}
A typical
signal-to-noise ratio (S/N) per wavelength bin of the spectra is about
15, and the heliocentric velocity correction was applied to the final
spectra.

\textcolor{red}{
In addition to the main targets in Table~\ref{tab:literatures}, we
also obtained the spectra of LHS~49 (on 2004-April-17) and HD~140538
(on 2004-May-06), which are the RV template and the RV standard
stars respectively.  These data were obtained using the same telescope
and the same instrument set up as for our main targets.  The S/N of
the both objects were about 60. 
}

@@

We have computed $p$ for all the objects (Table~\ref{tab:chisq_prob}),
and have plotted the results  
as a histogram of $-\log{p}$, shown in Fig.~\ref{fig:chisq-test}
(excluding the non-member USco~121; see explanation later). 
The figure clearly shows two distinctive populations: one on the left
(with small $-\log{p}$ values) occupied by the RV constant objects,
and one on the right (with the large $-\log{p}$ values)
occupied by the RV variable candidates.  The
expected distribution of $p$ computed consistently with our
uncertainty measurements is also shown in the same figure. The RV
constant population on the left side reasonably matches the expected
curve. 
\textcolor{red}{
To quantify this point, we have computed the
$\chi^{2}$ probability to test the goodness of the fit of the expected
distribution to the histogram. We restrict the fit to the first 4 bins
from the left $\left(0 < -\log_{10}p < 2 \right)$ the
histogram since the objects in the second bin from the right
$\left(2.5 < -\log_{10}p < 3 \right)$ are `potentially'
RV-variable object, although we have flagged them as the RV-constant
objects. In this analysis, we find the probability of 0.5 which
indicates that the fit is reasonable. Further, if we re-normalized the
expected distribution, accounting for that fact we neglected the
objects in the second bin from the right, we find the $\chi^{2}$
probability of the fit increases to 0.9.  This, in turn, indicates
that our uncertainties in RV values are reasonable, and an additional
systematic error may not be necessary in this analysis. 
}

\textcolor{red}{
There are two possible
source of systematic errors: (1)~the RV shifts caused by surface/atmospheric
activity, and (2)~the RV shifts due to the limit of the stability of
autoguiding of the telescope.  \citet{joergens:2006} found that (1) is
negligibly small for young VLM objects with $M_{*}<0.12\,M_{\sun}$.
The size of RV errors caused by (2) is expected to be $<1\,\kmps$
\citep{bailer-jones:2004}. Note that \citet{maxted:2005} found an
additional systematic error was not needed to be added to the RV
uncertainties of \citet{joergens:2006} who used the same instruments
as ours, and used the same method of estimating the
uncertainties.
}

@@

\caption{Histogram of the average heliocentric radial velocities of 14 UScoOB BD and VLM
objects listed in Table~\ref{tab:summary_results} (excluding USco~121,
a non-member). The gaussian fit (dashed) of the radial velocity
distribution gives a standard deviation of $1.0\,\kmps$ and the peak
position of $-6.3\,\kmps$. }
\label{fig:vrad_histrogram}

@@

The range of $v\,\sin i$ found among our objects is $3.6$--$55.6\,\kmps$,
and a similar range is also found by \citet{mohanty:2005}. 
Fig.~\ref{fig:vrot_histrogram} shows the histogram
of $v\,\sin i$ distribution for the UScoOB objects (14 objects excluding
USco 121, non member). The log-normal fit of this distribution gives
the peak position at $16.9\,\kmps$ with a standard deviation $\sigma=27.8\,\kmps$.
Using the $v\,\sin i$ data in \citet{joergens:2001}, the same histogram
bin size used for UScoOB objects and a log-normal fit, we find the
$v\,\sin i$ distribution of the BD and VLM stars (8 objects) in Cha~I
peaks at $15.4\,\kmps$, and has the standard deviation of $8.0\,\kmps$.
The peak of the distribution is similar to that of Upper Sco objects,
but the standard deviation of the distribution is much smaller
than that of our Upper Sco objects. The difference maybe due to the
very small sample. A similar fit was applied to the $v\,\sin i$
distribution of 14 higher mass T~Tauri stars in Cha~I using the data
of \citet{joergens:2001}, and we found a peak at $17.0\,\kmps$ with a
standard deviation $25.9\,\kmps$ which are very similar to those of
the Upper Sco brown dwarf candidates and VLM stars.

@@

\caption{Histogram of the rotational velocities ($v\,\sin{i}$) of 14 UScoOB BD and VLM
objects listed in Table~\ref{tab:summary_results} (excluding USco~121,
a non-member).  The log-normal fit (dashed) of the
rotational velocity distribution gives a standard deviation
$27.8\,\kmps$ and the peak position $16.9\,\kmps$.}
\label{fig:vrot_histrogram}

@@

In order to compute the detection probability as a function of
semimajor axis $a$, we take the following procedure: (1)~for each
object in our targets, we randomly select $10^5$ binaries using the
assumption stated above for a given value of $a$, and compute the RV
of the primary ($V_{1}$), (2)~evolve the orbit by the time separation
of the RV measurements used in the observations for this object, and
take another simulated measurement of RV ($V_{2}$), (3)~from $V_{1}$,
$V_{2}$ and the average uncertainty in RV ($\sigma_{\mathrm{RRV}}$)
from the observations, we compute the $\chi^{2}$ probability $p$,  and
flag the trial as a detection if $p<10^{-3}$ as done for the real
data, (4)~find the fraction of detections out of all random trials,
(5) repeat 1--4 for the range of $a$ between $10^{-3}$~and 10~au, and
(6) repeat 5 for the all targets, and find the detection probability
averaged over all targets as a function of $a$.

@@

An alternative explanation for the RV variations found in the binary
candidates above is a type of stellar pulsations found by
\citet{palla:2005}. They studied the non-adiabatic, linear
instability of very VLM stars and BD during the deuterium burning phase
in the core, and found unstable fundamental modes in the time-scale
between $\sim 1$ and $\sim 5$~h for the object mass from $0.02$ and
$0.1\,\Msun$.  By using their pulsation periods and by assuming that the
pulsation amplitudes is 5~per~cent of the radius of the objects, we
find that the expected RV variation can be $>6\,\kmps$.
As mentioned in Section~\ref{sec:observation-reduction}, for each object at a
given night, two separate spectra are obtained consecutively (with
time separations of $\sim 0.5$~h.). We have measured the RV from each
spectrum to check if there is a large jump in the RV measurements on a
short time-scale. The size of the shifts in the RV values measured in
the two consecutive spectra are found in between $0.02\pm0.10$
(USco~128) and  $2.04\pm1.67\,\kmps$ (USco~66), and the average size
of the shifts $0.36\,\kmps$, which is much smaller than the 
expected RV variation based on the models of \citet{palla:2005};
hence, we found little indication of the pulsations in our sample.

@@

Follow-up spectroscopic observations of the binary candidates
presented here are planned in near future. There are only a few RV
variable binary candidates identified in earlier surveys
(\citealt{guenther:2003}; \citealt{kenyon:2005}; \citealt{joergens:2006}).
As \citet{burgasser:2006} points out most of the current RV and
imaging surveys use samples from magnitude-limited survey, but one
should attempt to use the samples from volume-limited survey in order
to find a correct statistics on binary parameters more straightforwardly,
i.e. without correcting for bias. 
\textcolor{red}{
Finally, we note that the binary fraction found in this paper may be
over-estimated because of the nature of our survey (magnitude-limited
and incomplete), hence because of a possible bias towards the brighter
objects.
}

@@

We would like to thank the anonymous referee for constructive
comments and suggestions for improving the clarity of the manuscript.
We also thank the staff of VLT of the ESO for carrying out the observations
in service mode. RK is grateful for Rob Jeffries  and Tim Naylor for helpful
suggestions on the data analysis presented in this paper. We also
thank Matthew Bate for providing us valuable comments on
the manuscript. This work was supported by PPARC rolling grant
PP/C501609/1.

@@

\bibitem[{{Basri} \& {Reiners}(2006)}]{basri:2006}
{Basri} G., {Reiners} A., 2006, \aj, 132, 663

@@

% LocalWords:  BDs VLM OB Ophiuchi al Ryuichi Harries QL RV UScoOB BD echelle
% LocalWords:  VLT UVES RVs Cha bouy burgasser gizis kraus basri maxted PPl EEV
% LocalWords:  pinfield substellar duquennoy VLMs guenther joergens padoan CCDs
% LocalWords:  reipurth whitworth dezeeuw ardila luhman muzerolle Kueyen Parnal
% LocalWords:  LL ESO interorder autoguiding FWHM jones Oph wilking mohanty Sp
% LocalWords:  GY USco LHS HJD gaussian Cen garcia sanchez HD RRV pc CCF exp AJ
% LocalWords:  obs ApJ PASP MNRAS Chamaeleon semimajor siegler Praesepe stassun
% LocalWords:  mccaughrean segransan tinney palla kenyon Jeffries Naylor Bromm
% LocalWords:  natexlab Reiners Bonnell Moraux Brandner Baraffe ndez McElwain
% LocalWords:  astro ph de Zeeuw de Zeeuw de Bruijne Hoogerwerf Lub Blaauw
% LocalWords:  Battrick
